Efficient implementation of kernel filtering

### 2007-06-28 article/Article.tex
Acknowledgments
darcs-hash:20070625145545-2fc9d-2f35fef9e2ea60e348efd5fdc10d0e9d23673388.gz
make BoxAccumulateFunction take a functor argument
darcs-hash:20070627144902-2fc9d-5566cea487aa8d00bec8c879b4f7bf0599affaea.gz
add sigma filter - a test is still missing
darcs-hash:20070628103023-2fc9d-fd881259703ac3509863058b153c8a0fa7837cef.gz

\section{Mean and variance filters}
\label{sect:meanVar}
A simplified version of the moving histogram approach can also be used
to efficiently compute the mean or variance of an arbitrarily shaped
kernel. The histogram used for morphology and rank filters is a data
structure that, once established for one pixel, can be easily updated
to represent the kernel for a neighboring pixel. The data structure
required to perform the equivalent function for mean and variance
operations is much simpler -- all that is necessary is the sum of
pixel values under the kernel (or sum of squared values for variance
calculations), and the kernel size. The sum can be updated by adding
the new values and subtracting the old and the output computed by
dividing by the kernel size.

A more efficient strategy was proposed in \cite{Crow84} for
rectangular kernels. This approach first computes an accumulator image
in which each $i,j$ is replaced by the sum of voxels ``left'' and
``above'' it. This accumulator image may be computed recursively using
local neighborhood values. The mean of rectangles of any size may then
be computed using accumulated values at the rectangle corners. Figures
\ref{fig:accum} and \ref{fig:accumMean} illustrate the 2D case and the
update formulas are provided in Equations \ref{eq:accum} and
\ref{eq:accumMean}.

@@

\begin{equation}
\label{eq:accumMean}
m_{i,j} = \frac{1}{wh} (b_{i+w/2,j+h/2} + b_{i-w/2,j-h/2} - b_{i-w/2,j+h/2} + b_{i+w/2,j-h/2})
\end{equation}
where $m_{i,j}$ is the output of the mean filter at location $i,j$.

@@

>>>>

\section{Mean and standard deviation filters}
\label{sect:meanVar}
A simplified version of the moving histogram approach can also be used
to efficiently compute the mean or standard deviation of an arbitrarily shaped
kernel. The histogram used for morphology and rank filters is a data
structure that, once established for one pixel, can be easily updated
to represent the kernel for a neighboring pixel. The data structure
required to perform the equivalent function for mean and variance
operations is much simpler -- all that is necessary is the sum of
pixel values under the kernel (or sum of squared values for variance
calculations), and the kernel size. The sum can be updated by adding
the new values and subtracting the old and the output computed by
dividing by the kernel size.

A more efficient strategy was proposed in \cite{Crow84} for
rectangular kernels. This approach first computes an accumulator image
in which each $i,j$ is replaced by the sum of voxels ``left'' and
``above'' it. This accumulator image may be computed recursively using
local neighborhood values. The mean of rectangles of any size may then
be computed using accumulated values at the rectangle corners. Figures
\ref{fig:accum} and \ref{fig:accumMean} illustrate the 2D case and the
update formulas are provided in Equations \ref{eq:accum}, \ref{eq:accSum}, and
\ref{eq:accumMean}.

@@

\begin{equation}
\label{eq:accSum}
S_{i,j} = b_{i+w/2,j+h/2} + b_{i-w/2,j-h/2} - b_{i-w/2,j+h/2} + b_{i+w/2,j-h/2}
\end{equation}
where $S_{i,j}$ is the sum of the pixel values in the neighborhood centered at $i,j$.

\begin{equation}
\label{eq:accumMean}
m_{i,j} = \frac{1}{w.h}S_{i,j}
\end{equation}
where $m_{i,j}$ is the output of the mean filter at location $i,j$.

Using the same strategy, the standard deviation can be efficiently computed.
It requires an accumulator image of the squared pixels values, which can be
efficiently computed at the same time than the accumulator image. The update
formulas are provided in Equations \ref{eq:accum}, \ref{eq:accSum},
\ref{eq:accum2}, \ref{eq:accSum2} and \ref{eq:accumSigma}.

\begin{eqnarray}
\label{eq:accum2}
b2_{i,j} & = & \sum_{x \leq i, y \leq j} a_{x,y}^2 \\ \nonumber
	& = & b2_{i,j-1} + b2_{i-1, j} + a_{i,j}^2 - b2_{i-1, j-1}
\end{eqnarray}
where $b2_{i,j}$ is the accumulator value at location $i,j$ for the square image, and $a_{i,j}$ is the input image intensity at $i,j$.

\begin{equation}
\label{eq:accSum2}
S2_{i,j} = b2_{i+w/2,j+h/2} + b2_{i-w/2,j-h/2} - b2_{i-w/2,j+h/2} + b2_{i+w/2,j-h/2}
\end{equation}
where $S2_{i,j}$ is the sum of the squared pixel values in the neighborhood centered at $i,j$.

\begin{equation}
\label{eq:accumSigma}
\sigma_{i,j} = \sqrt{\frac{S2_{i,j} - \frac{S_{i,j}^2}{w.h}}{w.h-1}}
\end{equation}
where $\sigma_{i,j}$ is the output of the standard deviation filter at location $i,j$.

@@

\section{Acknowledgments}
We thank Dr Pierre Adenot and MIMA2 confocal facilities
(\url{http://mima2.jouy.inra.fr}) for providing the 3D test image.
